---
title: "Data preparation"
output:
  pdf_document: default
---

# Instructions

- You only need to submit the .Rmd of this file, not a PDF.

- You should __comment__ your code clearly to show what you've done to prepare the data.

- The purpose of this file is to use the data in the `data-raw` folder to create the data you will use in the report. The data you will use in the report should be saved in the `data` folder. It is good professional practice to make sure you're never directly modifying your raw data, but instead creating new datasets based on merges/manipulations that you need to reuse.

- Make sure you've taken a look at the hints for the web scraping and census API. 

- You may find the `write_rds()` function from the `readr` package helpful (it is loaded as part of the `tidyverse`).

- You do not need to keep the structure below.


# Set up

```{r, libraries}
#install packages if it errors on the concensus data and postal code data
#install.packages("sf")
install.packages("cancensus")
install.packages("haven")
install.packages("eeptools")

# Set up any libraries you need
library(tidyverse)
library(polite)
library(rvest)
library(janitor)
library(cancensus)
library(haven)
library(eeptools)
```

# Loading client data

```{r}
#loading raw rds files into environment
customer <- readRDS("data-raw/customer.Rds")
device <- readRDS("data-raw/device.Rds")
cust_dev <- readRDS("data-raw/cust_dev.Rds")
cust_sleep <- readRDS("data-raw/cust_sleep.Rds")
```

```{r}
mark = read_csv("data/marketing_data (1)")
```

# Getting external data

## Web scraping industry data

```{r}
#declaring the url
url <- "https://fitnesstrackerinfohub.netlify.app/"

#creating target, providing proper user agent
target <- bow(url,
              user_agent = "pablo.mercado@mail.utoronto.ca for STA303/1002 project",
              force = TRUE)

#scraping
html <- scrape(target)

#creating data set from web scrapping
device_data <- html %>% 
  html_elements("table") %>% 
  html_table() %>% 
  pluck(1)
```

# Census API
```{r}
options(cancensus.api_key = "CensusMapper_097b519443c084075f1669e5a98ed736",
        cancensus.cache_path = "cache") # this sets a folder for your cache


# get all regions as at the 2016 Census (2020 not up yet)
regions <- list_census_regions(dataset = "CA16")

regions_filtered <-  regions %>% 
  filter(level == "CSD") %>% # Figure out what CSD means in Census data
  as_census_region_list()

# We want to get household median income
census_data_csd <- get_census(dataset='CA16', regions = regions_filtered,
                          vectors=c("v_CA16_2397"), 
                          level='CSD', geo_format = "sf")

# Simplify to only needed variables
median_income <- census_data_csd %>% 
  as_tibble() %>% 
  select(CSDuid = GeoUID, contains("median"), Population) %>% 
  mutate(CSDuid = parse_number(CSDuid)) %>% 
  rename(hhld_median_inc = 2)
```
# Postal Code data
```{r}
#reading .sav file for postcode data
dataset = read_sav("data-raw/pccfNat_fccpNat_082021sav.sav")

#creating data frame and selecting only relevant variables
postcode <- dataset %>%
  select(PC, CSDuid)

```


## Cleaning the data
# joining the data WIP
```{r}
#creating a new data set joining all customer data by "cust_id"
cust_first = customer %>%
  right_join(cust_dev, by = "cust_id")

#creating a data set joining device data by "device_name"
device_full = device_data %>% 
  rename("device_name" = "Device name") %>%
  right_join(device, by = "device_name") %>%
  clean_names()

# get the device each customer owns
cust_mark = cust_first %>%
  left_join(device_full, by = "dev_id")

cust_full = cust_mark %>%
   right_join(cust_sleep, by ="cust_id")
```

# Modifying the sleep scores data for Modelling
```{r}
cust_full$emoji_modifier = fct_relevel(cust_full$emoji_modifier, "U+1F3FB")


cust_full = cust_full %>%
  mutate(age = floor(age_calc(dob, units = "years")),
         device_age = floor(age_calc(released_2, units = "years")))

cust_full = cust_full %>%
  filter(!is.na(sex))

cust_full = cust_full %>%
  mutate(age_group = case_when (
             age <= 20 ~ "Less than 20",
             age <= 30 ~ "20 to 30",
             age <= 40 ~ "30 to 40",
             age <= 50 ~ "40 to 50",
             age <= 60 ~ "50 to 60",
             age <= 70 ~ "60 to 70",
             age <= 80 ~ "70 to 80",
             age <= 100 ~ "80 or more",
             TRUE ~ "None"
             ))
```

# Loading client data


```{r}
marketing_data = cust_mark %>%
  select(-dev_id, -device_name,-line_2,-released_2)
```

```{r}
postcode = break_glass_in_case_of_emergency %>%
  rename(.,postcode = PC)

geo_data = postcode %>%
  left_join(median_income)
  
marketing_data = geo_data %>%
  left_join(postcode, by="postcode")
```

# cleaning the data
```{r}
marketing_data = marketing_data %>%
  drop_na(sex,pronouns)
  
marketing_data$emoji_modifier = recode(marketing_data$emoji_modifier, 
                                       `U+1F3FB` = "light skin tone",
                                       `U+1F3FC` = "medium-light skin tone",
                                       `U+1F3FD` = "medium skin tone",
                                       `U+1F3FE` = "medium-dark skin tone",
                                       `U+1F3FF` = "dark skin tone")

marketing_data$emoji_modifier = factor(marketing_data$emoji_modifier, levels = c("default skin tone",
                                                                                 "light skin tone","medium-light skin tone","medium skin tone","medium-dark skin tone","dark skin tone"))
marketing_data = marketing_data %>%
  mutate(affordable = ifelse(line == "Advance" | line == "Active", TRUE, FALSE),
         income_class = ifelse(hhld_median_inc <= 65327, "lower class",
                        ifelse(65327 < hhld_median_inc & hhld_median_inc <= 85981, "middle class", 
                        ifelse(hhld_median_inc > 85981, "upper class")))) 
```

```{r}
write.csv(marketing_data, "data/marketing_data")
```



# Model Comparisons for Sleep scores
```{r}
modrace2 = glmer(flags ~ emoji_modifier + age + (1 | cust_id), family='poisson', offset = log(duration), data= cust_full)
modrace = glmer(flags ~ emoji_modifier + sex + age + device_name + (1 | cust_id), 
        family='poisson', offset = log(duration), data= cust_full)
summary(modrace)
summary(modrace2)
# how do i justify  adding different variables to my model? Why is it showing this red output?
```

the mean flags per minute of sleep, controlling for each person is 2.165, 3.2414, 6.600, and 10.90 times higher for each darker skin tone than the lighest skin tone. Mingar needs to do something about this. 

```{r}
anova(modrace2, modrace, test="Chisq")

lmtest::lrtest(modrace2, modrace)
```


#exporting clean data to folder


```{r}
#  Should I justify why i have used this model?   Should i have the testing different models part in the main report? Yes, interpret the results and the summaries. where do i put the stuff about checking other models? in the actual report, but only state the final results. Should i state the assumptions for poisson in the main report? assumptiuons should be checked in the technical report.  should i do one mode with na, one model without, and then compare?
```

```{r}
# Finishing up data prep as informed by working on data summaries, visualizations, models and model assumption checking (and doing these things). should this all be in data prep section?
```


```{r}
# Our units/subjects are independent, even through observations within each subject are taken not to to be. (Replace subjects with ‘groups,’ same idea.) not sure if satisfied

# Random effects come from a normal distribution. not possible, our random effect is a factor

# The random effects errors and within-unit residual errors have constant variance. I.e., Are variances of data (transformed by the link function) homogeneous across categories? how to check this

# The chosen link function is appropriate / the model is correctly specified. check if the model fits poisson?
```

```{r}
# mean = variance
# should the main report include the process we took to reach the final model, or should it just have the report on the final model? Just final model. SHould we include info on the other variables? 
```


### Residual Plot
```{r}
# Do I put it in the main report
#yhatXresidZero
res.df <- data.frame(resid = residuals(modrace2), fit = fitted(modrace2))
ggplot(res.df, aes(x = fit, y = resid)) +
  geom_point() +
  ylab("Residuals from model") +
  xlab("Fitted values from model") +
  theme_minimal()
# why does the residual plot look weird? what does this say about the model? 
```




# Goodness of fit test
```{r}
1-pchisq(modrace$deviance, modrace$df.residual)
# how do i do gof for glmm?
```

```{r}
# limitations: there may be overdspersion ?
# NA's dropping from the model? 
```








# Loading client data
```{r}
marketing_data = cust_mark %>%
  select(-dev_id, -device_name,-line_2,-released_2)
```

```{r}
postcode = break_glass_in_case_of_emergency %>%
  rename(.,postcode = PC)
marketing_data = postcode %>%
  left_join(postcode, by="postcode")
```

# cleaning the data
```{r}
marketing_data = marketing_data %>%
  drop_na(sex,pronouns)
  
marketing_data$emoji_modifier = recode(marketing_data$emoji_modifier, 
                                       `U+1F3FB` = "light skin tone",
                                       `U+1F3FC` = "medium-light skin tone",
                                       `U+1F3FD` = "medium skin tone",
                                       `U+1F3FE` = "medium-dark skin tone",
                                       `U+1F3FF` = "dark skin tone")

marketing_data$emoji_modifier = factor(marketing_data$emoji_modifier, levels = c("default skin tone",
                                                                                 "light skin tone","medium-light skin tone","medium skin tone","medium-dark skin tone","dark skin tone"))
marketing_data = marketing_data %>%
  mutate(affordable = ifelse(line == "Advance" | line == "Active", TRUE, FALSE),
         income_class = ifelse(hhld_median_inc <= 65327, "lower class",
                        ifelse(65327 < hhld_median_inc & hhld_median_inc <= 85981, "middle class", 
                        ifelse(hhld_median_inc > 85981, "upper class")))) # why did you choose these criteria?
```

